---
title: "625 Sampling Project"
author: "Ellen Hickman, Mi Huynh, Maze Nead & Zupeng Zeng"
date: "2026-01-21"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readxl)
library(knitr)
library(ggrepel)
library(tidyverse)
options(scipen = 999)
```

## Download data
```{r}
schoolframe <- read.csv("https://raw.githubusercontent.com/ellenthickman/625sampling/refs/heads/main/MI_school_frame_head_counts.csv")
kable(schoolframe)
```
# Sampling Frame Analysis
## 1. Noncoverage
### Examine by: searching current # of MI schools (eg. charter)
### Prevalence:
### Solution: Acknowledge and accept in analysis. 

## 2. Blanks: Ellen.
### Examine by: “print NA value”
Missing District name ~ Non-public
### Prevalence: (eg. Frequency tables)
### Solution:

## 3. Duplicates: Mi. In the whole observation. 
### Examine by: “print unique”
### Prevalence:
### Solution:

## 4. Clustering: Zupeng.
### Examine by: 
### Prevalence:
### Solution:
```{r}
# Basic cleaning: ensure head-count columns are numeric (handles commas if present)
count_vars <- c("g7_totl","g8_totl","g9_totl","g10_totl","g11_totl","g12_totl","tot_all")

schoolframe <- schoolframe %>%
  mutate(across(all_of(count_vars),
                ~ as.numeric(gsub(",", "", as.character(.x)))))

# Sanity checks
N_rows     <- nrow(schoolframe)
N_bcode    <- n_distinct(schoolframe$BCODE)
N_district <- n_distinct(schoolframe$DCODE)

N_rows; N_bcode; N_district

# Check whether BCODE is unique (should be if each row is a unique school/building record)
bcode_dupes <- schoolframe %>%
  count(BCODE) %>%
  filter(n > 1) %>%
  arrange(desc(n))

bcode_dupes

# District-level school clustering
district_summary <- schoolframe %>%
  group_by(DCODE) %>%
  summarise(
    n_schools = n(),
    district_name_count = n_distinct(na.omit(District_Name)),
    total_enroll = sum(tot_all, na.rm = TRUE),
    .groups = "drop"
  )

# Distribution of schools per district
summary(district_summary$n_schools)

qs <- quantile(
  district_summary$n_schools,
  probs = c(.25, .50, .75, .90, .95, .99),
  na.rm = TRUE
)
qs

# Violin Plot
q_df <- tibble(
  prob  = names(qs),
  value = as.numeric(qs)
)

ggplot(district_summary, aes(x = "", y = n_schools)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_segment(
    data = q_df,
    aes(x = 0.85, xend = 1.15, y = value, yend = value),
    inherit.aes = FALSE,
    linewidth = 0.6
  ) +
  geom_text_repel(
    data = q_df,
    aes(x = 1.22, y = value, label = paste0(prob, ": ", round(value, 2))),
    inherit.aes = FALSE,
    direction = "y",
    nudge_x = 0.08,
    hjust = 0,
    size = 3,
    min.segment.length = 0
  ) +
  labs(
    title = "Distribution of Number of Schools by District",
    x = NULL,
    y = "Schools per district"
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    plot.margin = margin(5.5, 90, 5.5, 5.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

# Largest districts by number of schools
district_summary %>%
  arrange(desc(n_schools)) %>%
  slice_head(n = 10)

# Potential data issue: DCODE mapping to more than one District_Name
district_summary %>%
  filter(district_name_count > 1) %>%
  arrange(desc(n_schools))

# Duplicates of school name within a district
name_dupes_within_district <- schoolframe %>%
  filter(!is.na(District_Name), !is.na(BNAME)) %>%
  group_by(DCODE, District_Name, BNAME) %>%
  summarise(n_records = n(), .groups = "drop") %>%
  filter(n_records > 1) %>%
  arrange(desc(n_records))

name_dupes_within_district
```


## 5. Many-to-many matching: Maze.
### Examine by: 
### Prevalence:
### Solution: