---
title: "625 Sampling Project"
author: "Ellen Hickman, Mi Huynh, Maze Nead & Zupeng Zeng"
date: "2026-01-21"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readxl)
library(knitr)
library(ggrepel)
library(tidyverse)
options(scipen = 999)
```

## Download data
```{r}
schoolframe <- read.csv("https://raw.githubusercontent.com/ellenthickman/625sampling/refs/heads/main/MI_school_frame_head_counts.csv")
```

## Week 3: Sampling Frame Analysis
## 1. Noncoverage
### Examine by: searching current # of MI schools (eg. charter). See: https://www.mischooldata.org/student-enrollment-counts-report/
### Prevalence:
### Solution: Acknowledge and accept in analysis. 

## 2. Blanks: Ellen.
### Examine by: Missingness check. 
``` {r}
# a. Check missingness across variables as blank ""
sapply(schoolframe, function(x) sum(x == "", na.rm = TRUE))
```
Out of 2,443 schools in Michigan, 559 do not have listed District Names. Let's examine this prevalence by Non-Public/Public schools and by Region.

### Prevalence: (eg. Frequency tables)
```{r}
# b. Since missing district name might vary by public/private school status, let's check:
tab1 <- table(schoolframe$District_Name == "", schoolframe$Public.nonpublic); kable(tab1)
# c. By region, let's check:
tab2 <- table(schoolframe$District_Name == "", schoolframe$Region); kable(tab2)
tab3 <- table(factor(schoolframe$District_Name == "", levels = c(TRUE, FALSE)),
             factor(schoolframe$Region))
kable(prop.table(tab3, margin = 2))
# c. By district code, let's check:
tab4 <- table(schoolframe$DCODE, schoolframe$District_Name == "")
```
From Tables 1-3, we see that the majority of missing district name cases (547/559, 98%) belong to non-public (ie. private) schools, whereas only a handful of public schools are missing district names (12/1884, ~0.0%). It also looks like even though the private schools are missing District Names, they do belong to a District Code that other, public schools are also a part of. 

### Solution:
So, the solution to the problem is to 1) ignore District Names as a cluster variable, and rely on region or district code if needed. If need be, the district name could be 2) imputed from the dcode variable for reference in communicating the results. 

## 3. Duplicates: Mi. In the whole observation. 
### Examine by: “print unique”
### Prevalence:
### Solution:

## 4. Clustering: Zupeng.
### Examine by: 
### Prevalence:
### Solution:
```{r}
# Basic cleaning: ensure head-count columns are numeric (handles commas if present)
count_vars <- c("g7_totl","g8_totl","g9_totl","g10_totl","g11_totl","g12_totl","tot_all")

schoolframe <- schoolframe %>%
  mutate(across(all_of(count_vars),
                ~ as.numeric(gsub(",", "", as.character(.x)))))

# Sanity checks
N_rows     <- nrow(schoolframe)
N_bcode    <- n_distinct(schoolframe$BCODE)
N_district <- n_distinct(schoolframe$DCODE)

N_rows; N_bcode; N_district

# Check whether BCODE is unique (should be if each row is a unique school/building record)
bcode_dupes <- schoolframe %>%
  count(BCODE) %>%
  filter(n > 1) %>%
  arrange(desc(n))

bcode_dupes

# District-level school clustering
district_summary <- schoolframe %>%
  group_by(DCODE) %>%
  summarise(
    n_schools = n(),
    district_name_count = n_distinct(na.omit(District_Name)),
    total_enroll = sum(tot_all, na.rm = TRUE),
    .groups = "drop"
  )

# Distribution of schools per district
summary(district_summary$n_schools)

qs <- quantile(
  district_summary$n_schools,
  probs = c(.25, .50, .75, .90, .95, .99),
  na.rm = TRUE
)
qs

# Violin Plot
q_df <- tibble(
  prob  = names(qs),
  value = as.numeric(qs)
)

ggplot(district_summary, aes(x = "", y = n_schools)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_segment(
    data = q_df,
    aes(x = 0.85, xend = 1.15, y = value, yend = value),
    inherit.aes = FALSE,
    linewidth = 0.6
  ) +
  geom_text_repel(
    data = q_df,
    aes(x = 1.22, y = value, label = paste0(prob, ": ", round(value, 2))),
    inherit.aes = FALSE,
    direction = "y",
    nudge_x = 0.08,
    hjust = 0,
    size = 3,
    min.segment.length = 0
  ) +
  labs(
    title = "Distribution of Number of Schools by District",
    x = NULL,
    y = "Schools per district"
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    plot.margin = margin(5.5, 90, 5.5, 5.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

# Largest districts by number of schools
district_summary %>%
  arrange(desc(n_schools)) %>%
  slice_head(n = 10)

# Potential data issue: DCODE mapping to more than one District_Name
district_summary %>%
  filter(district_name_count > 1) %>%
  arrange(desc(n_schools))

# Duplicates of school name within a district
name_dupes_within_district <- schoolframe %>%
  filter(!is.na(District_Name), !is.na(BNAME)) %>%
  group_by(DCODE, District_Name, BNAME) %>%
  summarise(n_records = n(), .groups = "drop") %>%
  filter(n_records > 1) %>%
  arrange(desc(n_records))

name_dupes_within_district
```


## 5. Many-to-many matching: Maze.
### Examine by: 
### Prevalence:
### Solution:


## Week 4: SRS Sample Size
```{r}
cv_target <- 0.05

p_cig  <- 0.25  # ever smoked one cigarette
p_mj   <- 0.18  # ever smoked marijuana
mu_age <- 14    # age first approached
sd_age <- 10

# Element variances
S2_cig <- p_cig * (1 - p_cig)
S2_mj  <- p_mj  * (1 - p_mj)
S2_age <- sd_age^2

# Desired sampling variances
varstar_cig <- (cv_target * p_cig)^2
varstar_mj  <- (cv_target * p_mj)^2
varstar_age <- (cv_target * mu_age)^2

# Required n under SRS, ignoring FPC
n_cig <- S2_cig / varstar_cig
n_mj  <- S2_mj  / varstar_mj
n_age <- S2_age / varstar_age

# Round up to ensure CV <= target
n_req <- c(
  cig = ceiling(n_cig),
  mj  = ceiling(n_mj),
  age = ceiling(n_age)
)

# Summary
out <- data.frame(
  variable = c("Ever smoked one cigarette (p)", "Ever smoked marijuana (p)", "Age first approached (mean)"),
  expected_estimate = c(p_cig, p_mj, mu_age),
  element_variance_S2 = c(S2_cig, S2_mj, S2_age),
  target_CV = cv_target,
  desired_sampling_variance = c(varstar_cig, varstar_mj, varstar_age),
  required_n_SRS_ignore_FPC = n_req
)

out

```

## Week 5: Prepare a table of population counts to each of the regions
```{r}
schoolframe$tot_all <- as.numeric(schoolframe$tot_all)

regcounts <- schoolframe %>%
  as_tibble() %>%
  group_by(Region) %>%
  summarise(Count = sum(tot_all, na.rm = TRUE))
```

## Week 6: Interclass Correlation Coefficient (ICC) for clustering
```{r}
# Calculate Roh from Previous Study
n_prev <- 7500
a_prev <- 150
m_prev <- n_prev / a_prev

deff <- c(cig = 2.5, mj = 2.0, age = 1.7)

roh <- (deff - 1)/(m_prev - 1)

m_prev
roh
```

## Week 7: Optimal Sampling Plan
```{r}
budget <- 500000
cost_per_school <- 3000
cost_per_student <- 50

# Optimal m for each variable
m_opt_raw <- sqrt(cost_per_school * (1 - roh) / (cost_per_student * roh))
m_opt <- ceiling(m_opt_raw)   # round up to ensure CV target

# Corresponding optimal number of clusters a and total n
a_opt <- floor(budget / (cost_per_school + cost_per_student * m_opt))
n_opt <- a_opt * m_opt

opt_plan <- data.frame(
  variable = names(roh),
  roh = as.numeric(roh),
  m_opt_raw = as.numeric(m_opt_raw),
  m_opt = as.integer(m_opt),
  a_opt = as.integer(a_opt),
  n_opt = as.integer(n_opt),
  cost_used = as.numeric(a_opt * (cost_per_school + cost_per_student * m_opt))
)

opt_plan
```

```{r}
# Use the largest m_opt across variables to ensure CV target is met for all
m_overall <- max(m_opt)
m_overall
```

```{r}
# Corresponding a and n for overall m
a_overall <- floor(budget / (cost_per_school + cost_per_student * m_overall))
n_overall <- a_overall * m_overall

c(m_overall = m_overall, a_overall = a_overall, n_overall = n_overall,
  cost_used = a_overall * (cost_per_school + cost_per_student * m_overall))
```
```{r}
# New expected DEFFs under the chosen overall m
deff_new <- 1 + (m_overall - 1) * roh

# New expected SRS variances, ignoring FPC
var_srs_new <- c(
  cig = S2_cig / n_overall,
  mj  = S2_mj  / n_overall,
  age = S2_age / n_overall
)

# New expected sampling variances under cluster design
var_clust_new <- deff_new * var_srs_new

# CVs under new design, standard deviation/mean, expected mean values from week4 input
cv_new <- c(
  cig = sqrt(var_clust_new["cig"]) / p_cig,
  mj  = sqrt(var_clust_new["mj"])  / p_mj,
  age = sqrt(var_clust_new["age"]) / mu_age
)

# Compare to precision targets
varstar <- c(cig = varstar_cig, mj = varstar_mj, age = varstar_age)

precision_check <- data.frame(
  variable = c("cig","mj","age"),
  roh = as.numeric(roh),
  m_used = m_overall,
  a_used = a_overall,
  n_used = n_overall,
  deff_new = as.numeric(deff_new),
  S2 = c(S2_cig, S2_mj, S2_age),
  var_srs_new = as.numeric(var_srs_new),
  var_clust_new = as.numeric(var_clust_new),
  var_target = as.numeric(varstar),
  meets_var_target = as.logical(var_clust_new <= varstar),
  cv_new = as.numeric(cv_new),
  meets_cv_target = as.logical(cv_new <= cv_target)
)

precision_check
```

```{r}
# Calculate overall sample fraction
N_pop <- 830138
f_overall <- n_overall / N_pop
f_overall
```
- Since the f is small, <0.05, we can keep excluding the FPC in variance calculations for it would have little impact.

### Looping over each variable's optimal allocation
```{r}
# Store results
precision_tables <- list()

for(i in 1:nrow(opt_plan)) {
  
  # Extract optimal design values for this variable
  m_overall <- opt_plan$m_opt[i]
  a_overall <- opt_plan$a_opt[i]
  n_overall <- opt_plan$n_opt[i]
  
  # New expected DEFFs under this m
  deff_new <- 1 + (m_overall - 1) * roh
  
  # New expected SRS variances
  var_srs_new <- c(
    cig = S2_cig / n_overall,
    mj  = S2_mj  / n_overall,
    age = S2_age / n_overall
  )
  
  # Cluster design variances
  var_clust_new <- deff_new * var_srs_new
  
  # CVs under new design
  cv_new <- c(
    cig = sqrt(var_clust_new["cig"]) / p_cig,
    mj  = sqrt(var_clust_new["mj"])  / p_mj,
    age = sqrt(var_clust_new["age"]) / mu_age
  )
  
  # Targets
  varstar <- c(cig = varstar_cig, mj = varstar_mj, age = varstar_age)
  
  # Precision check table
  precision_tables[[ opt_plan$variable[i] ]] <- data.frame(
    variable = c("cig","mj","age"),
    roh = as.numeric(roh),
    m_used = m_overall,
    a_used = a_overall,
    n_used = n_overall,
    deff_new = as.numeric(deff_new),
    S2 = c(S2_cig, S2_mj, S2_age),
    var_srs_new = as.numeric(var_srs_new),
    var_clust_new = as.numeric(var_clust_new),
    var_target = as.numeric(varstar),
    meets_var_target = as.logical(var_clust_new <= varstar),
    cv_new = as.numeric(cv_new),
    meets_cv_target = as.logical(cv_new <= cv_target)
  )
}

precision_tables$cig <- precision_tables$cig %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))
precision_tables$mj <- precision_tables$mj %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))
precision_tables$age <- precision_tables$age %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))

# View tables
precision_tables$cig
precision_tables$mj
precision_tables$age

# a is the number of schools
# m is the number of clusters
# 9 strata
# As clusters increase, variance also increase because homogeneity increases
# A lower CV is better because there's less variation between samples
```

## Week 8
1. First-stage sampling fractions for each stratum, where the total number of schools to sample at the first stage in each stratum is defined by your proportionate allocation of the $n_opt$ clusters.
A. Proportionate allocation
$$
\frac{N_h}{N} = W_h
$$
$$
n_h = n * \frac{N_h}{N} = n * W_h
$$
B. Input known values
```{r}
n <- 4698 # Total overall n
m <- 54 # No. of schools (clusters)
a <- 87 # No. of students within school (cluster size)
regcounts[1:2]

#Calculate stratum weight
regcounts$w_h <- regcounts$Count/sum(regcounts$Count)
```
C. Calculate total number of schools to sample at the first stage in each stratum (m)
```{r}
regcounts$m_h_raw <- regcounts$w_h * m
regcounts$m_h_round <- round(regcounts$m_h_raw); sum(regcounts$m_h_round)

m - sum(regcounts$m_h_round) # Number of schools left to allocate after rounding

# Initialize
regcounts$m_h <- regcounts$m_h_round

#Add 1 to the lowest three strata from the largest three strata to achieve 9 strata
regcounts$m_h[regcounts$m_h == 0] <- regcounts$m_h[regcounts$m_h == 0] + 1

# Subtract 1 from region 9
regcounts$m_h[regcounts$Region == 9] <- regcounts$m_h[regcounts$Region == 9] - 1 

# Paired
regcounts$m_h_pair <- c(2,2,2,2,2,2,10,10,22) #Evenly allocated to avoid singleton problem. Noting Strata 6 is now underrepresented compared to adjusted strata #1-5

sum(regcounts$m_h)
sum(regcounts$m_h_pair)

# Proportionate allocation of schools across strata
regcounts
```
D. Calculate total number of schools to sample at the first stage in each stratum (n) (EXTRA)
```{r}
#Calculate sample size within each stratum
regcounts$n_raw <- (n*regcounts$w_h); regcounts

#Round down strata #6 since f_i is closest to .50 threshold & Round to nearest integer for all else
regcounts <- regcounts %>%
  mutate(
    n_raw = (n * Count) / sum(Count),
    n_h = ifelse(Region == 6, floor(n_raw), round(n_raw))
  )

sum(regcounts$n_h); regcounts
```

2. Sort the list to achieve implicit stratification within the regions, using tot_all from previous year as MOS
A. Sort the list
```{r}
schoolframe$tot_all <- as.numeric(gsub(",", "", as.character(schoolframe$tot_all)))

schoolframe_sorted <- schoolframe %>%
  arrange(Region, tot_all)
```

3. Compute your zone size (Fractional intervals) for systematic PPeS sampling within each of the 9 strata
A. Zone size
$$
zone\ size = \frac{\sum_{i\epsilon h}{MOS_{hi}}}{n_h}
$$
B. Calculate 
```{r}
regcounts$zone_size <- sum(regcounts$Count) / regcounts$n_h; regcounts
```

4. Select the allocated number of schools within each stratum using PPeS sampling using systematic selection based on the fractional intervals (Zone size) 
```{r}
# Step 5: PPeS systematic selection within each stratum
library(sampling)
selected_schools <- list()

for(r in unique(schoolframe_sorted$Region)) {
  
  frame_r <- schoolframe_sorted %>%
    filter(Region == r)
  
  m_h_r <- regcounts$m_h_pair[regcounts$Region == r]
  M_h_r <- sum(frame_r$tot_all, na.rm = TRUE)
  
  # Skip if no schools allocated to this region
  if(length(m_h_r) == 0 || m_h_r == 0) next
  
  # PPeS inclusion probabilities within stratum
  pik <- m_h_r * frame_r$tot_all / M_h_r
  
  # to be safe: cap at 1
  pik[pik > 1] <- 1
  
  # systematic PPeS selection
  s <- UPsystematic(pik = pik)
  
  # if nothing selected, skip safely
  if(sum(s) == 0) next
  
  ss_r <- getdata(frame_r, s)
  ss_r$pik <- pik[s == 1]
  ss_r$selected_region <- r
  
  selected_schools[[as.character(r)]] <- ss_r
}

selected_schools_df <- bind_rows(selected_schools)

selected_schools_df
table(selected_schools_df$Region)


# Check whether the number selected matches allocation!!
selected_counts <- selected_schools_df %>%
  group_by(Region) %>%
  summarise(selected_schools = n(), .groups = "drop")

regcounts %>%
  select(Region, m_h_pair) %>%
  left_join(selected_counts, by = "Region")
```
5. Report first-stage sampling fraction (f) within each of the nine strata under PPes
A. First stage sampling fraction
$$
f_i = \frac{MOS_i}{zone\ size}
$$
B. Calculate
```{r}
regcounts$f_i <- regcounts$Count / regcounts$zone_size
```
C. Report: First-stage sampling fraction within each stratum under PPeS
```{r}
# Stratum-level first-stage sampling fraction
regcounts %>%
  select(Region, Count, m_h_pair, zone_size, f_i) # stratum-level first-stage fraction: f_h^ = m_h/m_h

# School-level PPeS first-stage inclusion probabilities
selected_schools_df %>%
  select(Region, BCODE, tot_all, pik) # school-level first-stage fraction: pik = m_h * tot_all / M_h
```

6. Form pseudo-strata to meet paired selection model to estimate sampling variance
```{r}
selected_schools_df <- selected_schools_df %>%
  arrange(Region, tot_all) %>%   # keep the PPS-selected schools in sorted order within region
  group_by(Region) %>%
  mutate(
    order_in_region = row_number(),
    pseudo_stratum_within_region = ceiling(order_in_region / 2) # create pseudo-strata of size 2 within each region for paired variance estimation
  ) %>%
  ungroup()

# Unique pseudo-stratum ID across regions
selected_schools_df <- selected_schools_df %>%
  mutate(
    pseudo_stratum_id = paste0("R", Region, "_P", pseudo_stratum_within_region)
  )
```

## Week 10
1. For each of the nine strata, compute the required number of students to subsample from each sampled school based on the stratified PPeS design in order to maintain epsem across all (9) strata.
```{r}
# Step 1: Required number of students per sampled school to maintain epsem

# number of students to sample per school in each stratum to maintain epsem
regcounts$b_h_raw <- regcounts$n_h / regcounts$m_h_pair
# Round up to ensure we meet the required number of students per stratum
regcounts$b_h <- ceiling(regcounts$b_h_raw)

regcounts %>%
  select(Region, n_h, m_h_pair, b_h_raw, b_h)
```

2. Do each of the schools that you sampled in a given region have the minimum sufficient size, given the stratum-specific subsample sizes computed in Task #1 (Week 10)? Do subsequent schools on the list have the minimum sufficient size? If not, what will you do?
```{r}
# Step 2: Check if sampled schools have sufficient size
selected_schools_df <- selected_schools_df %>%
  left_join(regcounts %>% select(Region, b_h), by = "Region") %>%
  mutate(
    # check if the total number of students in the school meets the required number to sample
    sufficient_size = tot_all >= b_h
  )

selected_schools_df %>%
  select(Region, BCODE, tot_all, b_h, sufficient_size)

# Identify schools that do not meet the sufficient size requirement
selected_schools_df %>%
  filter(!sufficient_size) %>%
  select(Region, BCODE, tot_all, b_h)
```

3. Begin to describe how you will physically select the subsample of students within a given sampled school (or set of linked schools). What will your second-stage sampling rate be for a given school within a given stratum? How will you acquire the updated rosters from each school? What technique will you use to select the sample at the specified second-stage Rate?
```{r}
# Second stage sampling rate
# f / f_i 
regcounts$f_h <- (regcounts$m_h*regcounts$n_h)/regcounts$Count
# Overall f for each stratum (f = m_h*n_h)/MOS_h
regcounts$subsample_rate <- regcounts$f_h/regcounts$f_i
regcounts$m_star <- regcounts$subsample_rate*regcounts$Count
```


4. Write down the overall sampling fraction based on the stratified PPeS design, indicating the overall probability of inclusion for a given student, from a given school (or linked set of schools), in a given stratum. Be careful with notation. Keep in mind that the MOS (Measure of Size) values used for the sampled schools at the first stage and the denominator at the second stage (Did you sample a single school? Or a linked set of schools?) will depend on your response to Task #2 above.
```{r}
overall_f <- sum(regcounts$f_h)
# overall_f <- (54*4698)/830138
```

## Week 11
### Download data
```{r}
studentlist <- read.csv("https://raw.githubusercontent.com/ellenthickman/625sampling/main/sample_school_student_list.csv")
head(studentlist)
```

1. Find subsampling rate of region 7
$$
f_{j|i} = \frac{m^*_h}{MOS_h}
$$
$$
x_i = f_{j|i} * M_h
$$
a. Input known values
```{r}
mstar_7 = 12
MOS_7 = 242

M_7 = 219 #Actual size of school. To find x_7
```
b. Calculate
```{r}
fji_7 <- mstar_7 / MOS_7; fji_7

x_7 <- fji_7 * M_7; x_7
```
2. Find two seperate probabilities
$$
x_7 = x_1(1-p)+x_2(p)
$$
a. Fractional interval (inverse of subsampling rate)
$$
\frac{1}{f_{j|i}}
$$

```{r}
p <- 0.8595

x_7_compare <- 11*(p)+10*(1-p)
x_7; x_7_compare #Verify
```
3. Report
```{r}
1/fji_7; p; 1-p
```
With a fractional interval 20.17, select a sample of 11 employees with
probability 0.86 or a sample of 10 employees with probability 0.14.

3. Select 
The sample is sorted by grade (7->8) then grouped by teacher (random)

a. Proportion of students within each grade
```{r}
table(studentlist$Grade)
```


### RECYCLE BIN ###
A. Overall sampling rate (Lec. 8 PPeS, p. 58)
$$
f = \frac{ma*}{\sum_{i\epsilon h}MOS_{hi}}; where\\ 
m = \#\ of\ clusters\ and\\
a* = targeted\ sub\ sample\ size
$$
B. Calculate overall sampling fraction
```{r}
#Calculate overall sampling rate
f <- (m*a)/sum(regcounts$Count); f
```
C. Strata sampling rate (Lec. 8 PPeS, p. 58)
$$
f = \frac{m_hm_h*}{\sum_{i\epsilon h}MOS_{hi}}; where\\ 
m = \#\ of\ clusters\ and\\
a* = targeted\ sub\ sample\ size
$$
D. Calculate
```{r}
regcounts$f_h <- (regcounts$m_h * regcounts$m_h) / sum(regcounts$Count); regcounts
```

D. Sampling rate of clusters at the first stage in each stratum (n)
$$
n_h = \frac{nMOS_i}{\sum_{i}MOS_{i}}
$$

D. Sampling rate of clusters (schools) at the first stage in each stratum (m)
$$
m_h = \frac{mMOS_h}{\sum_{i\epsilon h}MOS_{h}}
$$
F. f_h
$$
f_h = \frac{n_hm_h^*}{\sum_{i\epsilon h}MOS_{hi}}
$$

