---
title: "625 Sampling Project"
author: "Ellen Hickman, Mi Huynh, Maze Nead & Zupeng Zeng"
date: "2026-01-21"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readxl)
library(knitr)
library(ggrepel)
library(tidyverse)
options(scipen = 999)
```

## Download data
```{r}
schoolframe <- read.csv("https://raw.githubusercontent.com/ellenthickman/625sampling/refs/heads/main/MI_school_frame_head_counts.csv")
kable(schoolframe)
```

## Week 3: Sampling Frame Analysis
## 1. Noncoverage
### Examine by: searching current # of MI schools (eg. charter). See: https://www.mischooldata.org/student-enrollment-counts-report/
### Prevalence:
### Solution: Acknowledge and accept in analysis. 

## 2. Blanks: Ellen.
### Examine by: Missingness check. 
``` {r}
# a. Check missingness across variables as blank ""
sapply(schoolframe, function(x) sum(x == "", na.rm = TRUE))
```
Out of 2,443 schools in Michigan, 559 do not have listed District Names. Let's examine this prevalence by Non-Public/Public schools and by Region.

### Prevalence: (eg. Frequency tables)
```{r}
# b. Since missing district name might vary by public/private school status, let's check:
tab1 <- table(schoolframe$District_Name == "", schoolframe$Public.nonpublic); kable(tab1)
# c. By region, let's check:
tab2 <- table(schoolframe$District_Name == "", schoolframe$Region); kable(tab2)
tab3 <- table(factor(schoolframe$District_Name == "", levels = c(TRUE, FALSE)),
             factor(schoolframe$Region))
kable(prop.table(tab3, margin = 2))
# c. By district code, let's check:
tab4 <- table(schoolframe$DCODE, schoolframe$District_Name == ""); kable(tab4)
```
From Tables 1-3, we see that the majority of missing district name cases (547/559, 98%) belong to non-public (ie. private) schools, whereas only a handful of public schools are missing district names (12/1884, ~0.0%). It also looks like even though the private schools are missing District Names, they do belong to a District Code that other, public schools are also a part of. 

### Solution:
So, the solution to the problem is to 1) ignore District Names as a cluster variable, and rely on region or district code if needed. If need be, the district name could be 2) imputed from the dcode variable for reference in communicating the results. 

## 3. Duplicates: Mi. In the whole observation. 
### Examine by: “print unique”
### Prevalence:
### Solution:

## 4. Clustering: Zupeng.
### Examine by: 
### Prevalence:
### Solution:
```{r}
# Basic cleaning: ensure head-count columns are numeric (handles commas if present)
count_vars <- c("g7_totl","g8_totl","g9_totl","g10_totl","g11_totl","g12_totl","tot_all")

schoolframe <- schoolframe %>%
  mutate(across(all_of(count_vars),
                ~ as.numeric(gsub(",", "", as.character(.x)))))

# Sanity checks
N_rows     <- nrow(schoolframe)
N_bcode    <- n_distinct(schoolframe$BCODE)
N_district <- n_distinct(schoolframe$DCODE)

N_rows; N_bcode; N_district

# Check whether BCODE is unique (should be if each row is a unique school/building record)
bcode_dupes <- schoolframe %>%
  count(BCODE) %>%
  filter(n > 1) %>%
  arrange(desc(n))

bcode_dupes

# District-level school clustering
district_summary <- schoolframe %>%
  group_by(DCODE) %>%
  summarise(
    n_schools = n(),
    district_name_count = n_distinct(na.omit(District_Name)),
    total_enroll = sum(tot_all, na.rm = TRUE),
    .groups = "drop"
  )

# Distribution of schools per district
summary(district_summary$n_schools)

qs <- quantile(
  district_summary$n_schools,
  probs = c(.25, .50, .75, .90, .95, .99),
  na.rm = TRUE
)
qs

# Violin Plot
q_df <- tibble(
  prob  = names(qs),
  value = as.numeric(qs)
)

ggplot(district_summary, aes(x = "", y = n_schools)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_segment(
    data = q_df,
    aes(x = 0.85, xend = 1.15, y = value, yend = value),
    inherit.aes = FALSE,
    linewidth = 0.6
  ) +
  geom_text_repel(
    data = q_df,
    aes(x = 1.22, y = value, label = paste0(prob, ": ", round(value, 2))),
    inherit.aes = FALSE,
    direction = "y",
    nudge_x = 0.08,
    hjust = 0,
    size = 3,
    min.segment.length = 0
  ) +
  labs(
    title = "Distribution of Number of Schools by District",
    x = NULL,
    y = "Schools per district"
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    plot.margin = margin(5.5, 90, 5.5, 5.5),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

# Largest districts by number of schools
district_summary %>%
  arrange(desc(n_schools)) %>%
  slice_head(n = 10)

# Potential data issue: DCODE mapping to more than one District_Name
district_summary %>%
  filter(district_name_count > 1) %>%
  arrange(desc(n_schools))

# Duplicates of school name within a district
name_dupes_within_district <- schoolframe %>%
  filter(!is.na(District_Name), !is.na(BNAME)) %>%
  group_by(DCODE, District_Name, BNAME) %>%
  summarise(n_records = n(), .groups = "drop") %>%
  filter(n_records > 1) %>%
  arrange(desc(n_records))

name_dupes_within_district
```


## 5. Many-to-many matching: Maze.
### Examine by: 
### Prevalence:
### Solution:


## Week 4: SRS Sample Size
```{r}
cv_target <- 0.05

p_cig  <- 0.25  # ever smoked one cigarette
p_mj   <- 0.18  # ever smoked marijuana
mu_age <- 14    # age first approached
sd_age <- 10

# Element variances
S2_cig <- p_cig * (1 - p_cig)
S2_mj  <- p_mj  * (1 - p_mj)
S2_age <- sd_age^2

# Desired sampling variances
varstar_cig <- (cv_target * p_cig)^2
varstar_mj  <- (cv_target * p_mj)^2
varstar_age <- (cv_target * mu_age)^2

# Required n under SRS, ignoring FPC
n_cig <- S2_cig / varstar_cig
n_mj  <- S2_mj  / varstar_mj
n_age <- S2_age / varstar_age

# Round up to ensure CV <= target
n_req <- c(
  cig = ceiling(n_cig),
  mj  = ceiling(n_mj),
  age = ceiling(n_age)
)

# Summary
out <- data.frame(
  variable = c("Ever smoked one cigarette (p)", "Ever smoked marijuana (p)", "Age first approached (mean)"),
  expected_estimate = c(p_cig, p_mj, mu_age),
  element_variance_S2 = c(S2_cig, S2_mj, S2_age),
  target_CV = cv_target,
  desired_sampling_variance = c(varstar_cig, varstar_mj, varstar_age),
  required_n_SRS_ignore_FPC = n_req
)

out

```

## Week 5: Prepare a table of population counts to each of the regions
```{r}
schoolframe$tot_all <- as.numeric(schoolframe$tot_all)

regcounts <- schoolframe %>%
  as_tibble() %>%
  group_by(Region) %>%
  summarise(Count = sum(tot_all, na.rm = TRUE))
```

## Week 6: Interclass Correlation Coefficient (ICC) for clustering
```{r}
# Calculate Roh from Previous Study
n_prev <- 7500
a_prev <- 150
m_prev <- n_prev / a_prev

deff <- c(cig = 2.5, mj = 2.0, age = 1.7)

roh <- (deff - 1)/(m_prev - 1)

m_prev
roh
```

## Week 7: Optimal Sampling Plan
```{r}
budget <- 500000
cost_per_school <- 3000
cost_per_student <- 50

# Optimal m for each variable
m_opt_raw <- sqrt(cost_per_school * (1 - roh) / (cost_per_student * roh))
m_opt <- ceiling(m_opt_raw)   # round up to ensure CV target

# Corresponding optimal number of clusters a and total n
a_opt <- floor(budget / (cost_per_school + cost_per_student * m_opt))
n_opt <- a_opt * m_opt

opt_plan <- data.frame(
  variable = names(roh),
  roh = as.numeric(roh),
  m_opt_raw = as.numeric(m_opt_raw),
  m_opt = as.integer(m_opt),
  a_opt = as.integer(a_opt),
  n_opt = as.integer(n_opt),
  cost_used = as.numeric(a_opt * (cost_per_school + cost_per_student * m_opt))
)

opt_plan
```

```{r}
# Use the largest m_opt across variables to ensure CV target is met for all
m_overall <- max(m_opt)
m_overall
```

```{r}
# Corresponding a and n for overall m
a_overall <- floor(budget / (cost_per_school + cost_per_student * m_overall))
n_overall <- a_overall * m_overall

c(m_overall = m_overall, a_overall = a_overall, n_overall = n_overall,
  cost_used = a_overall * (cost_per_school + cost_per_student * m_overall))
```
```{r}
# New expected DEFFs under the chosen overall m
deff_new <- 1 + (m_overall - 1) * roh

# New expected SRS variances, ignoring FPC
var_srs_new <- c(
  cig = S2_cig / n_overall,
  mj  = S2_mj  / n_overall,
  age = S2_age / n_overall
)

# New expected sampling variances under cluster design
var_clust_new <- deff_new * var_srs_new

# CVs under new design, standard deviation/mean, expected mean values from week4 input
cv_new <- c(
  cig = sqrt(var_clust_new["cig"]) / p_cig,
  mj  = sqrt(var_clust_new["mj"])  / p_mj,
  age = sqrt(var_clust_new["age"]) / mu_age
)

# Compare to precision targets
varstar <- c(cig = varstar_cig, mj = varstar_mj, age = varstar_age)

precision_check <- data.frame(
  variable = c("cig","mj","age"),
  roh = as.numeric(roh),
  m_used = m_overall,
  a_used = a_overall,
  n_used = n_overall,
  deff_new = as.numeric(deff_new),
  S2 = c(S2_cig, S2_mj, S2_age),
  var_srs_new = as.numeric(var_srs_new),
  var_clust_new = as.numeric(var_clust_new),
  var_target = as.numeric(varstar),
  meets_var_target = as.logical(var_clust_new <= varstar),
  cv_new = as.numeric(cv_new),
  meets_cv_target = as.logical(cv_new <= cv_target)
)

precision_check
```

```{r}
# Calculate overall sample fraction
N_pop <- 830138
f_overall <- n_overall / N_pop
f_overall
```
- Since the f is small, <0.05, we can keep excluding the FPC in variance calculations for it would have little impact.

### Looping over each variable's optimal allocation
```{r}
# Store results
precision_tables <- list()

for(i in 1:nrow(opt_plan)) {
  
  # Extract optimal design values for this variable
  m_overall <- opt_plan$m_opt[i]
  a_overall <- opt_plan$a_opt[i]
  n_overall <- opt_plan$n_opt[i]
  
  # New expected DEFFs under this m
  deff_new <- 1 + (m_overall - 1) * roh
  
  # New expected SRS variances
  var_srs_new <- c(
    cig = S2_cig / n_overall,
    mj  = S2_mj  / n_overall,
    age = S2_age / n_overall
  )
  
  # Cluster design variances
  var_clust_new <- deff_new * var_srs_new
  
  # CVs under new design
  cv_new <- c(
    cig = sqrt(var_clust_new["cig"]) / p_cig,
    mj  = sqrt(var_clust_new["mj"])  / p_mj,
    age = sqrt(var_clust_new["age"]) / mu_age
  )
  
  # Targets
  varstar <- c(cig = varstar_cig, mj = varstar_mj, age = varstar_age)
  
  # Precision check table
  precision_tables[[ opt_plan$variable[i] ]] <- data.frame(
    variable = c("cig","mj","age"),
    roh = as.numeric(roh),
    m_used = m_overall,
    a_used = a_overall,
    n_used = n_overall,
    deff_new = as.numeric(deff_new),
    S2 = c(S2_cig, S2_mj, S2_age),
    var_srs_new = as.numeric(var_srs_new),
    var_clust_new = as.numeric(var_clust_new),
    var_target = as.numeric(varstar),
    meets_var_target = as.logical(var_clust_new <= varstar),
    cv_new = as.numeric(cv_new),
    meets_cv_target = as.logical(cv_new <= cv_target)
  )
}

precision_tables$cig <- precision_tables$cig %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))
precision_tables$mj <- precision_tables$mj %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))
precision_tables$age <- precision_tables$age %>% 
  mutate(cost_used = as.numeric(a_used * (cost_per_school + cost_per_student * m_used)))

# View tables
precision_tables$cig
precision_tables$mj
precision_tables$age

# a is the number of schools
# m is the number of clusters
# 9 strata
# As clusters increase, variance also increase because homogeneity increases
# A lower CV is better because there's less variation between samples
```

## Week 8
1. First-stage sampling fractions (f) for each stratum (proportional allocation of the 63 schools)
2. PPeS selection of schools for each 9 strata
3. Sort the list to achieve implicit stratification within the regions, using tot_all as MOS
4. Compute your zone size for systematic PPeS sampling within each of the 9 strata
5. Select the allocated number of schools within each stratum using PPeS sampling using systematic selection based on the fractional intervals (Actual size)
6. First-stage sampling fraction (f) within each of the nine strata under PPes
7. Form pseudo-strata to meet paired selection model to estimate sampling variance

```{r}
# Proportional allocation
N_allocate <- sum(regcounts$Count)
n_used <- 4698
m_used <- 54
a_used <- 87

# Step 1: Allocate number of schools across strata
regcounts$weight <- regcounts$Count / N_allocate
regcounts$a_h_raw <- regcounts$weight * a_used
regcounts$a_h_floor <- floor(regcounts$a_h_raw)
regcounts$frac_a <- regcounts$a_h_raw - regcounts$a_h_floor

r_a <- a_used - sum(regcounts$a_h_floor) # Number of schools left to allocate after flooring

regcounts <- regcounts %>%
  arrange(desc(frac_a)) %>%
  mutate(add_a = ifelse(row_number() <= r_a, 1, 0)) %>% # Add 1 to the top r_a strata with the largest fractional parts
  arrange(Region)

regcounts$a_h <- regcounts$a_h_floor + regcounts$add_a  # Final allocated number of schools per stratum after adjusting for rounding
sum(regcounts$a_h)

# Proportionate allocation of schools across strata
regcounts

# Step 2: Allocate total students across strata
regcounts$n_h_raw <- regcounts$weight * n_used
regcounts$n_h_floor <- floor(regcounts$n_h_raw)
regcounts$frac_n <- regcounts$n_h_raw - regcounts$n_h_floor

r_n <- n_used - sum(regcounts$n_h_floor)

regcounts <- regcounts %>%
  arrange(desc(frac_n)) %>%
  mutate(add_n = ifelse(row_number() <= r_n, 1, 0)) %>%
  arrange(Region)

regcounts$n_h <- regcounts$n_h_floor + regcounts$add_n
sum(regcounts$n_h)

# Proportionate allocation of students across strata
regcounts

# Proportionate allocation is the number of students in each strata if we choose n=4698 from the precision tables

# PPeS
# Probabilities proportionate to estimated size
f <- n_used/sum(schoolframe$tot_all) # 4698 / total of all elements 

# Step 3: Sort the list
schoolframe$tot_all <- as.numeric(gsub(",", "", as.character(schoolframe$tot_all)))

schoolframe_sorted <- schoolframe %>%
  arrange(Region, tot_all)

# Step 4: Zone size
school_counts <- schoolframe_sorted %>%
  as_tibble() %>%
  group_by(Region) %>%
  summarise(
    A_h = n(),                          # number of schools in stratum
    M_h = sum(tot_all, na.rm = TRUE),   # total MOS in stratum
    .groups = "drop"
  )

regcounts <- regcounts %>%
  left_join(school_counts, by = "Region")

regcounts$zone_size <- regcounts$M_h / regcounts$a_h

# First-stage school sampling fraction in each stratum
regcounts$f1_h <- regcounts$a_h / regcounts$A_h

regcounts

# Step 5: PPeS systematic selection within each stratum
library(sampling)

selected_schools <- list()

for(r in unique(schoolframe_sorted$Region)) {
  
  frame_r <- schoolframe_sorted %>%
    filter(Region == r)
  
  a_h_r <- regcounts$a_h[regcounts$Region == r]
  M_h_r <- sum(frame_r$tot_all, na.rm = TRUE)
  
  # Skip if no schools allocated to this region
  if(length(a_h_r) == 0 || a_h_r == 0) next
  
  # PPeS inclusion probabilities within stratum
  pik <- a_h_r * frame_r$tot_all / M_h_r
  
  # to be safe: cap at 1
  pik[pik > 1] <- 1
  
  # systematic PPeS selection
  s <- UPsystematic(pik = pik)
  
  # if nothing selected, skip safely
  if(sum(s) == 0) next
  
  ss_r <- getdata(frame_r, s)
  ss_r$pik <- pik[s == 1]
  ss_r$selected_region <- r
  
  selected_schools[[as.character(r)]] <- ss_r
}

selected_schools_df <- bind_rows(selected_schools)

selected_schools_df
table(selected_schools_df$Region)


# Check whether the number selected matches allocation!!
selected_counts <- selected_schools_df %>%
  group_by(Region) %>%
  summarise(selected_schools = n(), .groups = "drop")

regcounts %>%
  select(Region, a_h) %>%
  left_join(selected_counts, by = "Region")


# Step 6: First-stage sampling fraction within each stratum under PPeS
# Stratum-level first-stage sampling fraction
regcounts %>%
  select(Region, A_h, a_h, f1_h, M_h, zone_size) # stratum-level first-stage fraction: f_h^ = a_h/A_h

# School-level PPeS first-stage inclusion probabilities
selected_schools_df %>%
  select(Region, BCODE, tot_all, pik) # school-level first-stage fraction: pik = a_h * tot_all / M_h


# Step 7: Pseudo-strata for paired variance estimation
selected_schools_df <- selected_schools_df %>% # sort by region and MOS
  arrange(Region, tot_all) %>% # create pseudo-strata by pairing schools within each region
  group_by(Region) %>%
  mutate(
    order_in_region = row_number(), # order schools within region by MOS
     # each pair of schools gets the same pseudo-stratum ID within the region
    pseudo_stratum = ceiling(order_in_region / 2) 
  ) %>%
  ungroup()

selected_schools_df %>%
  select(Region, BCODE, tot_all, pseudo_stratum)
```

## Week 10
1. For each of the nine strata, compute the required number of students to subsample from each sampled school based on the stratified PPeS design in order to maintain epsem across all (9) strata.
2. Do each of the schools that you sampled in a given region have the minimum sufficient size, given the stratum-specific subsample sizes computed in Task #1 (Week 10)? Do subsequent schools on the list have the minimum sufficient size? If not, what will you do?
3. Begin to describe how you will physically select the subsample of students within a given sampled school (or set of linked schools). What will your second-stage sampling rate be for a given school within a given stratum? How will you acquire the updated rosters from each school? What technique will you use to select the sample at the specified second-stage Rate?
4. Write down the overall sampling fraction based on the stratified PPeS design, indicating the overall probability of inclusion for a given student, from a given school (or linked set of schools), in a given stratum. Be careful with notation. Keep in mind that the MOS (Measure of Size) values used for the sampled schools at the first stage and the denominator at the second stage (Did you sample a single school? Or a linked set of schools?) will depend on your response to Task #2 above.

```{r}
# Step 1: Required number of students per sampled school to maintain epsem

# number of students to sample per school in each stratum to maintain epsem
regcounts$b_h_raw <- regcounts$n_h / regcounts$a_h
# Round up to ensure we meet the required number of students per stratum
regcounts$b_h <- ceiling(regcounts$b_h_raw)

regcounts %>%
  select(Region, n_h, a_h, b_h_raw, b_h)

# Step 2: Check if sampled schools have sufficient size
selected_schools_df <- selected_schools_df %>%
  left_join(regcounts %>% select(Region, b_h), by = "Region") %>%
  mutate(
    # check if the total number of students in the school meets the required number to sample
    sufficient_size = tot_all >= b_h
  )

selected_schools_df %>%
  select(Region, BCODE, tot_all, b_h, sufficient_size)

# Identify schools that do not meet the sufficient size requirement
selected_schools_df %>%
  filter(!sufficient_size) %>%
  select(Region, BCODE, tot_all, b_h)
```